## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(BAS)
```

### Load data


```{r load-data}
load("movies.Rdata")
```

* * *

## Part 1: Data
This dataset is randomly selected from IMDB and Rotten Tomatoes, containing information about 651 movies produced before 2016. Since the sampling size is large enough and less than 10% of the population, plus the sampling was random, we can generalize these results to all movies produced before 2016 in these two websites. 

As there is no random assignment performed, the study is observational and hence no conclusions about causality can be made.Instead, the results could be used to demonstrate correlation.
* * *

## Part 2: Data manipulation
We are going to create required features, feature_film, drama, mpaa_rating_R, oscar_season and summer_season.


Feature_film: movies that are feature films:
```{r}
movies$feature_film<-ifelse(movies$title_type == 'Feature Film',c('yes'),c('no'))
```

Drama: movies whose genre is drama:
```{r}
movies$drama<-ifelse(movies$genre == 'Drama',c('yes'),c('no'))
```

mpaa_rating_R: movies that are R rated:
```{r}
movies$mpaa_rating_R<-ifelse(movies$mpaa_rating== 'R',c('yes'),c('no'))
```

oscar_season: movies that are released from October to December:
```{r}
movies$oscar_season<-ifelse(movies$thtr_rel_month %in% 10:12, 'yes','no')
```

summer_season: movies that are released from May to August:
```{r}
movies$summer_season<-ifelse(movies$thtr_rel_month %in% 5:8, 'yes','no')
```

* * *

## Part 3: Exploratory data analysis

First of all, we examine our response variable of interest audience_score:

```{r}
summary(movies$audience_score)
```

As shown above, the mean audience_score is 62.35 and the median is 65. 

```{r}
ggplot(data=movies, aes (x = audience_score)) + geom_histogram(binwidth=5)
```

It appears that the data is slightly left skewed. 

Now we perform an analysis of the relationships between audience score and newly featured variables by creating boxplots and credible interval.

1. the relatinship bewween audience_score and feature_film:
```{r}
boxplot(audience_score~feature_film, data = movies)
```

It appears that feature films have relatively lower audience score.

```{r}
bayes_inference(y=audience_score,x=feature_film, data = movies, statistic = 'mean', type = 'ci')
```
We can see the credible interval proves that there is indeed difference existing between these two grouops. 

For other variables, we conduct the same analysis. 

2. the relationship between audience score and the genre of the movie.

```{r}
boxplot(audience_score~drama, data = movies)
```

```{r}
bayes_inference(y=audience_score,x=drama, data = movies, statistic = 'mean', type = 'ci')
```

The average score of drama films are slightly higher than other genres, and credible interval proves the difference between two groups.


3. the relationship between audience score and mpaa rating.
```{r}
boxplot(audience_score~mpaa_rating_R, data = movies)
```

```{r}
bayes_inference(y=audience_score,x=mpaa_rating_R, data = movies, statistic = 'mean', type = 'ci')
```

The plots demonstrate that the difference between R rated and non-R rated films is not significant. 


4. the relationship between audience score and the time of the movie released.
```{r}
boxplot(audience_score~oscar_season, data = movies)
```

```{r}
bayes_inference(y=audience_score,x=oscar_season, data = movies, statistic = 'mean', type = 'ci')
```

The difference in the average audience score for movies released in oscar season and those who are not is small, and the credible interval contains 0.

```{r}
boxplot(audience_score~summer_season, data = movies)
```

```{r}
bayes_inference(y=audience_score,x=summer_season, data = movies, statistic = 'mean', type = 'ci')
```

There is almost no difference in the average audience between films released during the summer season and those who are not. The credible interval contains 0. 

* * *

## Part 4: Modeling

First of all, we remove N/A in the dataset.
```{r}
data_movies = na.omit(movies)
```

To build a model we will use all the requested variables and Bayesian Model Averaging(BMA) method to find the optimal model and optimal variables. We choose the ZS-null as a priorto assign equal weight to all of the models. 

```{r}
movies_bma <- bas.lm(audience_score~ feature_film + drama + runtime + mpaa_rating_R + thtr_rel_year + oscar_season + summer_season + imdb_rating + imdb_num_votes + critics_score + best_pic_nom + best_pic_win + best_actor_win + best_actress_win + best_dir_win + top200_box, data = data_movies, prior = "ZS-null", modelprior = uniform(),method = 'BAS')
summary(movies_bma)
```


For model diagnostic, We now examine the residuals:

```{r}
plot(movies_bma, which = 1)
```

The residuals appear to be randomly distributed around 0. For low BMA predictions, some resuals are larger and observations 126, 216 and 251 appear to be outliers. 

```{r}
plot(movies_bma, which = 4)
```


As the inclusion plot shows, the optimal model has 4 coefficients: the intercept, the runtime, tne IMDB rating and the critics score. They all have a probability of inclusion greater than 0.5. 

Now we visualize the model rank:
```{r}
image(movies_bma, rotate = F)
```

IMDB rating and critic scores are the most widely used variables.


Now let us look at the coefficients.
```{r}
coef <- coefficients(movies_bma)
confint(coef)
```

To address our research question, the coefficients' credible intervals shows that among 3 variables which are included in the optimal model, IMDB rating and critics scores are positively associated with the audience score.
* * *

## Part 5: Prediction

The movie we choose to predict is Arrival, which was released on 2016. The data for Arrival is from 
http://www.imdb.com/title/tt2543164/?ref_=nv_sr_1
https://www.rottentomatoes.com/m/arrival_2016
The audience score is 82, the critics score is 94, and IMDB rating is 7.9

```{r}
arrival <- data.frame(imdb_rating = 7.9, critics_score = 94)
fit <- bas.lm(audience_score~imdb_rating + critics_score, data = movies)
movie_2016 <- predict(fit,newdata = arrival, interval = 'confidence')
movie_2016$fit
```

The predicted audience score is 85.6, not very different the actual score 82.


* * *

## Part 6: Conclusion
This report goes through the process of creating a bayesian regression model that can be used to find out what factors make a movie popular. Using BMA method, the two variables for the optimal model are critic scores and IMDB rating. In the prediction section, our model did a great job. We have noticed that there are a few outliers in the lower bound of audience score, so there is still some space to make the model more accurate. 
